---
title: "Model Prediction"
author: "Marieke"
date: "October 28, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE)
```

## Loading Library

```{r}
library(adehabitat)
library(lubridate)
library(data.table)
library(raster)
library(rgdal)
library(rhdf5)

pro=CRS("+init=epsg:28992")
WGS84<-CRS("+init=epsg:4326")

starttime<-as.POSIXct("2016-06-27")
stoptime<-as.POSIXct("2016-06-30")

time.seq<-seq(from=starttime,to=stoptime,by="day")

kleur.breaks<-seq(0,500,by=10)
kleur.cols<-terrain.colors(length(kleur.breaks-1))

savedir<-"/nobackup/users/dirksen/radiation/Rdata/MLPredictions/"
```

## Natural earth data for plotting
Natural earth data is open source data, these small files are easilly projected. We create the object "fun" which will be used for the plotting routine of the raster stack. 
```{r}
#Natural Earth dataset: unprojected shape files
mymap.unpro=readOGR(dsn='Rdata/NaturalEarthData/ne_10m_admin_0_countries',layer="ne_10m_admin_0_countries") # Read in (unprojected) map data
mymap.pro=spTransform(mymap.unpro, pro) # Reproject the map

mymap.unpro_lakes=readOGR(dsn='Rdata/NaturalEarthData/ne_10m_lakes',layer="ne_10m_lakes") # Read in (unprojected) map data
mymap.pro_lakes=spTransform(mymap.unpro_lakes, pro) # Reproject the map

fun <- function() {
  plot(mymap.pro,add=TRUE)
  plot(mymap.pro_lakes,add=TRUE)
}
```

## Models
Here we load or models stored in the folder Rdata/MLmodels. ctree and treebag are large models and take a long time to run, if you want to add these models uncomment the lines. 
```{r}
list.files("Rdata/MLmodels/")

#ctree <- readRDS("Rdata/MLmodels/ctree.rds")
cubist <- readRDS("Rdata/MLmodels/cubist.rds")
earth <- readRDS("Rdata/MLmodels/earth.rds")
#fuzzy <- readRDS("Rdata/MLmodels/fuzzy.rds")
gbm <- readRDS("Rdata/MLmodels/gbm.rds")
glm <- readRDS("Rdata/MLmodels/glm.rds")
knn <- readRDS("Rdata/MLmodels/knn.rds")
lm <- readRDS("Rdata/MLmodels/lm.rds")
# treebag <- readRDS("Rdata/MLmodels/treebag.rds")

```

## Data Processing and Analysis
Based on the time period that was selected the script searches for the satellite grids from that specific date. The satellite grid is created from the hdf5 file and then, note that due to the scan direction of the satellite we transpose the irradiance and flip the grid. The satellite grid is cropped to the extent of the Netherlands. To the raster stack we add the layers of the different predictions and the object as .rds.  
```{r pressure, echo=FALSE}
for (i in 1:length(time.seq)){
t<-time.seq[i]
Y<-year(t)
M<-month(t)

if (M<=9) {
  M<-paste0(0,M)
} else {print(M)}

D<-mday(t)

if (D<=9) {
  D<-paste0(0,D)
} else {print(D)}

file.loc<-"/net/pc150398/nobackup_1/users/meirink/siccs_wouter/SICCS/daymean/harmonie_proj/year"
file.loc<-gsub("year",Y,file.loc)
files.in.loc<-list.files(file.loc,pattern=".h5")

d.sat<-gsub("daymean_reproj_","",files.in.loc)
d.sat<-gsub(".h5","",d.sat)
I=which(d.sat==paste0(Y,M,D))
if (length(I)==0){
  print("NO FILE")
  # out<-data.frame(time,"no file")
  # names(out)<-NULL
  # write.table(out,file="/nobackup/users/dirksen/Radiation_Obs_Satellite/output/statistical_summary/NOdata_days.csv",
  #             append=TRUE,sep=",",row.names=FALSE)
} else {
file.sat<-"/net/pc150398/nobackup_1/users/meirink/siccs_wouter/SICCS/daymean/harmonie_proj/year/daymean_reproj_yyyymmdd.h5"
file.sat<-gsub("year",Y,file.sat)
file.sat<-gsub("yyyymmdd",paste0(Y,M,D),file.sat)

h5ls(file.sat)
data.direct.irradiance<-h5read(file.sat,"direct irradiance")
data.diffuse.irradiance<-h5read(file.sat,"diffuse irradiance")
data.total.irradiance<-data.direct.irradiance+data.diffuse.irradiance
data.total.irradiance<-t(data.total.irradiance)

data.lat<-h5read(file.sat,"/lat") #Latitude
data.lon<-h5read(file.sat,"/lon") #Longitude

data.lat[which(data.lat==-999)]<-NA # replace no data value with NA (this case -999)
data.lon[which(data.lon==-999)]<-NA # replace no data value with NA (this case -999)

r<-raster(data.total.irradiance,crs=WGS84,
xmn=min(data.lon,na.rm=T),
xmx=max(data.lon,na.rm=T),
ymn=min(data.lat,na.rm=T),
ymx=max(data.lat,na.rm=T))
rr<-flip(r,direction='y')
rr<-projectRaster(rr,crs=pro)

r.NED<-crop(rr,extent(12621.630033977,278621.630033977,305583.0457758,620583.0457758))
names(r.NED)<-"SAT"
r.NED.matrix<-as.matrix(r.NED)
if (length(which(is.na(r.NED.matrix)))>2000) {
  print("NO data")
  # out<-data.frame(time,"no data NL")
  # names(out)<-NULL
  # write.table(out,file="/nobackup/users/dirksen/Radiation_Obs_Satellite/output/statistical_summary/NOdata_days.csv",
  #             append=TRUE,sep=",",row.names=FALSE)
} else {

st<-stack(r.NED)
st$day<-yday(t)
st$knn<-raster::predict(model=knn,object=st)
st$earth<-raster::predict(model=earth,object=st)
st$gbm<-raster::predict(model=gbm,object=st)
st$glm<-raster::predict(model=glm,object=st)
st$lm<-raster::predict(model=lm,object=st)
st$cubist<-raster::predict(model=cubist,object=st)

st<-dropLayer(st,2)

savefile <- paste0(savedir,paste0(Y,M,D),".rds")
saveRDS(st,savefile)


plot(st,col=kleur.cols,breaks=kleur.breaks,legend=FALSE,addfun=fun)
}
}
}
```


