---
title: "Documentation"
author: "Marieke"
date: "August 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Pre-processing Satellite data
## Library path and loading librarys
```{r}
.libPaths("/usr/people/dirksen/R-20160725/x86_64-redhat-linux-gnu-library/3.3/")

#librarys
library(rhdf5)
library(maptools)
library(raster)
library(ncdf4)
library(adehabitat)
library(rgdal)
```

## Initialization 
* define files location, load Rdata
* projection for the save files RD-coordinates
* extract time from names .h5 files
```{r}
load("/nobackup/users/dirksen/Radiation_Obs_Satellite/auxcillary_data/sun.Rdata") #created with sunriset function for Cabauw station
data(wrld_simpl)
files_dir<-"/nobackup/users/dirksen/Radiation_Obs_Satellite/Satellite_data/"

#setwd("/nobackup/users/dirksen/Radiation_Obs_Satellite/satellite_meirink_world/") #data satellite world

# initialization
pro=CRS("+init=epsg:28992")

#files
files<-list.files(pattern=".h5",full.names = TRUE)

#For data from Jan-Fokke
#Onbewolkte dag: 18,23 juli 2014
format<-"msg3_yyyymmdd_hhmm_00000_fl_cpp.h5"
time<-gsub("msg3_","",files)
time<-gsub("_00000_fl_cpp.h5","",time)

##create save file for sunrise and sunset times
t<-as.POSIXct(time,format="%Y%m%d_%H%M",tz="GMT")
day<-as.POSIXct((as.Date(t)))
```

## Processing all the files
* all the files between sunrise and sunset 
* from the .nc files a raster file is created
* saved to .ascii [NOTE: not ideal, an .nc file does have metadata]
```{r}
#as an example 2 files are selected
for (i in 40:41){
d<-day[i]
d<-format(d,format="%Y-%m-%d")
sun_up_down<-sun[which(format(sun$sunrise,"%Y-%m-%d")==d),]
if  (t[i]>sun_up_down$sunrise & t[i]<sun_up_down$sunset){
# myfile<-tempfile()
file<-files[i]
#print(time[i])

#find the data structure and variables
data.str<-h5ls(file)

data.sds<-h5read(file,"/sds") #the data

data.att<-h5readAttributes(file,"/sds") #description of the data (description, gain, intercept, no_data_value)
print(data.att)

data.sds<-data.sds*data.att$gain #note: the intergers should be multiplied with the gain! in case of sds gain=0.1
data.sds[which(data.sds==as.numeric(data.att$no_data_value*data.att$gain))]<-NA # replace no data value with NA (this case -1)

data.sds<-t(data.sds)

#loading the metadata for the field
#note: output from the h5 file is a matrix, not easy to cbind
data.att<-h5readAttributes(file,"/lat")

data.lat<-h5read(file,"/lat") #Latitude
data.lon<-h5read(file,"/lon") #Longitude

data.lat[which(data.lat==as.numeric(data.att$no_data_value))]<-NA # replace no data value with NA (this case -1)
data.lon[which(data.lon==as.numeric(data.att$no_data_value))]<-NA # replace no data value with NA (this case -1)

data.lat<-data.lat*data.att$gain #from interger to numeric
data.lon<-data.lon*data.att$gain

#create a raster file with coordinates (reference system: WGS84)
r<-raster(data.sds,
          xmn=min(data.lon,na.rm=T),
          xmx=max(data.lon,na.rm=T),
          ymn=min(data.lat,na.rm=T),
          ymx=max(data.lat,na.rm=T),
          crs=CRS("+init=epsg:4326"))
          #crs=CRS("+init=epsg:4326"))
#
#Visualization of the raster file
# plot.new()
# plot(r,main=time[i],asp=1)
# plot(wrld_simpl,add=TRUE)

#Reproject and crop the raster file
r<-projectRaster(r,crs=pro)

r2<-crop(r,extent(12621.630033977,278621.630033977,305583.0457758,620583.0457758))

#Save file as ASCII
GRID_name<-gsub(".h5","",file)
GRID_new<-paste0(files_dir,GRID_name)

writeRaster(r2, GRID_new, format="ascii",overwrite=TRUE)
} else
  print("night")
}

```

# Pre-processing HARM38 

## Initialization
```{r}
setwd("/nobackup/users/dirksen/Radiation_Obs_Satellite/HARMONIE38/NC")
files_dir<-"/nobackup/users/dirksen/Radiation_Obs_Satellite/HARMONIE38/ASCII/"
allfiles<-list.files(pattern=".nc")
nodatafiles<-list.files(pattern="\\_00000")

ncfiles<-allfiles[!allfiles %in% nodatafiles]
nl.grd <- read.asciigrid("/nobackup/users/dirksen/GIS/nl_clip_1km.txt")
projection(nl.grd)<-pro
nl.ras<-raster(nl.grd)
```

## Data Pre-processing
* from the .nc files a raster file is created
* saved to .ascii [NOTE: not ideal, an .nc file does have metadata]
* data is accumulative!
```{r}

for (i in 1:length(ncfiles)){
 print(i)
    file<-ncfiles[i]
    b<-brick(file,var='aswsn')
    r<-raster(b,layer=1)
    projection(r)<-CRS("+init=epsg:4326")
    
    #Reproject and crop the raster file
    r<-projectRaster(r,crs=pro)
    r2<-crop(r,extent(12621.630033977,278621.630033977,305583.0457758,620583.0457758)) 
    r2.1km<-resample(r2,nl.ras)
    
    #Save file as ASCII
    GRID_name<-gsub(".nc","",file)
    GRID_new<-paste0(files_dir,GRID_name)

    writeRaster(r2.1km, GRID_new, format="ascii",overwrite=TRUE)
    }
```

## From accumulative data to time averaged
* Datum is obtained from file name
* The first 24 hours of the run are used (as the run starts at nigh there is enough spin-up time for the model)
* Calculating time averaged values ([asc2-asc1]/[d2-d1]), where d is the time in seconds
* Save as .rda
```{r}
path<-"/nobackup/users/dirksen/Radiation_Obs_Satellite/HARMONIE38/ASCII/"

lst.dates<-list.files(path=path,pattern='.asc$',full.names=FALSE)
lst<-list.files(path=path,pattern='.asc$',full.names=TRUE)

#extracting the time from the file name
runstart<-gsub("HA38_N25_","",lst.dates)
runstart<-gsub("0000_0","",runstart)
runstart<-gsub(".asc","",runstart)

d<-as.POSIXct(runstart,format="%Y%m%d%H%M")
#d<-sort(d)
indices<-which(!is.na(d)) #True==NA value
d<-d[indices]
#we only want the files from the first 24 hours, so now select
lst<-lst[indices]

#test is equal
print(length(d))
print(length(lst))

save.file<-"/nobackup/users/dirksen/Radiation_Obs_Satellite/HARMONIE38/rda/rdata.rda"

for (i in 1:(length(d)-1)) {
fname<-gsub("rdata",runstart[i+1],save.file)
  
lst.1<-lst[i]
lst.2<-lst[i+1]

lst.asc1<-import.asc(lst.1)
lst.asc2<-import.asc(lst.2)
d.1<-d[i]
d.2<-d[i+1]

out.2<-(lst.asc2-lst.asc1)/(as.numeric(d.2)-as.numeric(d.1))

plot(out.2,main=d.2)

save(out.2,d.2,file=fname)

}
```