---
title: "Radiation and cloud cover"
author: "Marieke"
date: "February 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE)
```
## Introduction
Does cloud cover change the differences between the solar radiation products for the ground based measurements and the satellite measurements?

## Loading library
```{r library}
.libPaths("/usr/people/dirksen/R-20160725/x86_64-redhat-linux-gnu-library/3.3/")
library(data.table)
library(lubridate)
library(mgcv)
library(adehabitat)
library(automap)
library(caret)
library(doParallel)
library(kernlab)
library(reshape)
library(raster)
library(rgdal)
library(SDMTools)
library(plyr)
library(corrplot)
library(ggplot2)
pro=CRS("+init=epsg:28992")
save_dir<-"/nobackup/users/dirksen/radiation/Rdata/MLmodels/Linear_models/"
```


## Comparing the 2 products
To compare the satellite data with the point measurements the R function over is used. The data was stored in the file "sat_over_obs.csv". Using the data.table package and mgcv package we made a summary of the results. We find that the differences between the two products are small. There is a yearly pattern in the data: during winter months the satellite derived products has slightly higher values (max around 3W/m2) while in summer the opposit is true. 

To start with, a simple linear model is built using the ground based measurements of solar irradiance to predict the satellite radiation product. 
```{r observations}
obs.df <- fread("Rdata/sat_over_obs2.csv")
names(obs.df) <- c("date","DS_CODE","Q","x","y","SAT","DIFF")
obs.df$date <- as.Date(obs.df$date)
obs.df$day <- yday(obs.df$date)
obs.df$STN<-gsub("_H","",obs.df$DS_CODE)
setkey(obs.df,date)

#plot for each individual station
ggplot(data=obs.df,aes(x=date,y=DIFF)) +
  geom_point(size=1) +
  facet_wrap(~DS_CODE)

#Q(1) What is the mean difference/bias for each station?
#Q(2) Are station 215 and 323 the same stations?

# obs.df<-subset(obs.df,select=c("Q","SAT"))
obs.df <- obs.df[complete.cases(obs.df),]


ggplot(data=obs.df,aes(x=Q,y=SAT))+
  geom_point()


```

```{r cloud cover}
cloudcover<-readRDS("Rdata/cloudcover.rda")
cloudcover$DATE<-as.Date(cloudcover$DATE)

obs.Q.NG<-merge(obs.df,cloudcover,by.x=c("STN","date"),by.y=c("STN","DATE"))

ggplot(data=obs.Q.NG,aes(x=Q,y=SAT))+
  geom_point(aes(colour=factor(NG)))

q50<-aggregate(DIFF~NG,obs.Q.NG,FUN='quantile',probs=0.5)
q50$DIFF<-round(q50$DIFF,2)

ggplot(data=obs.Q.NG,aes(NG,DIFF)) +
  geom_boxplot() +
  coord_cartesian(ylim=c(-10,10)) +
  # ylim(-10,10) + # fit does change! Also boxplots do change!
  geom_smooth(method="loess",se=FALSE,color="blue",aes(group=1)) +
  geom_text(data=q50,aes(label=DIFF,y=DIFF-0.85)) +
  xlab("cloud cover [octa]") +
  ylab("difference [W/m2]") 
  
  


```


```{r sun}
sunshine<-readRDS("Rdata/sunshine.rda")

obs.Q.S<-merge(obs.df,sunshine,by.x=c("STN","date"),by.y=c("DS_CODE","IT_DATETIME"))

ggplot(data=obs.Q.S,aes(x=Q,y=SAT))+
  geom_point(aes(colour=factor(REH1.SQ24)))

ggplot(data=obs.Q.S,aes(REH1.SQ24,DIFF)) +
  geom_boxplot() +
  coord_cartesian(ylim=c(-10,10)) +
  # ylim(-10,10) + # fit does change! Also boxplots do change!
  geom_smooth(method="loess",se=FALSE,color="blue",aes(group=1)) +
  # geom_text(data=q50,aes(label=DIFF,y=DIFF-0.85)) +
  xlab("sunshine duration [hours]") +
  ylab("difference [W/m2]") 
```
```{r snow}
snow<-readRDS("/nobackup/users/dirksen/radiation/Rdata/snow.rda")
snow$DATE<-as.Date(snow$DATE,format="%Y%m%d")

snowobs<-merge(obs.df,snow,by.x=c("STN","date"),by.y=c("DS_CODE","DATE"))

snowobs<-snowobs[which(snowobs$SNOW!= "997" & snowobs$SNOW!= "998" & snowobs$SNOW!= "999"),]
snowobs$SNOW.num<-as.numeric(snowobs$SNOW)
setkey(snowobs,SNOW.num)

q50<-aggregate(DIFF~SNOW,snowobs,FUN='quantile',probs=0.5)
q50$DIFF<-round(q50$DIFF,2)

ggplot(data=snowobs,aes(SNOW.num,DIFF)) +
  geom_point() +
  # coord_cartesian(ylim=c(-10,10)) +
  # ylim(-10,10) + # fit does change! Also boxplots do change!
  # geom_smooth(method="glm",se=FALSE,color="blue",aes(group=1)) +
  # geom_text(data=q50,aes(label=DIFF,y=DIFF-2),size=2.5) +
  stat_smooth() +
  xlab("snow [cm]") +
  ylab("difference [W/m2]") 
```



